### 2024-05-26 getting-started.qmd
20240526 - additional data management

```{r}
nfl_players$position_group == "RB"
```

@@

```{r}
nfl_players$position_group != "RB"
```

@@

```{r}
nfl_players$weight > 300
```

@@

```{r}
nfl_players$weight < 300
```

@@

```{r}
nfl_players$weight >= 300
```

@@

```{r}
nfl_players$weight <= 300
```

@@

```{r}
nfl_players$position_group %in% c("QB","RB","WR")
```

@@

```{r}
!(nfl_players$position_group %in% c("QB","RB","WR"))
```

@@

```{r}
is.na(nfl_players$college_name)
```

@@

```{r}
!is.na(nfl_players$college_name)
```

@@

```{r}
nfl_players$position_group == "RB" & nfl_players$weight > 230
```

@@

```{r}
nfl_players$position_group == "RB" | nfl_players$weight > 230
```

@@

```{r, eval = FALSE}
dataframe[rowsToKeep, columnsToKeep]
```

@@

```{r}
nfl_players$display_name
```

@@

```{r}
nfl_players$display_name[which(nfl_players$position_group == "RB")]
```

@@

>>>>

nfl_players$position_group == "RB"
```

@@

nfl_players$position_group != "RB"
```

@@

nfl_players$weight > 300
```

@@

nfl_players$weight < 300
```

@@

nfl_players$weight >= 300
```

@@

nfl_players$weight <= 300
```

@@

nfl_players$position_group %in% c("QB","RB","WR")
```

@@

!(nfl_players$position_group %in% c("QB","RB","WR"))
```

@@

is.na(nfl_players$college_name)
```

@@

!is.na(nfl_players$college_name)
```

@@

nfl_players$position_group == "RB" & nfl_players$weight > 230
```

@@

nfl_players$position_group == "RB" | nfl_players$weight > 230
```

@@

dataframe[rowsToKeep, columnsToKeep]
```

@@

nfl_players$display_name
```

@@

nfl_players$display_name[which(nfl_players$position_group == "RB")]
```

@@

## View Data {#sec-viewData}

### All Data {#sec-viewDataAll}

To view data, use the following syntax:

View(nfl_players)
```

### First 6 Rows/Elements {#sec-viewDataHead}

To view only the first six rows (if a data frame) or elements (if a vector), use the following syntax:

```{r}
head(nfl_players)
head(nfl_players$display_name)
```

## Data Characteristics {#sec-dataCharacteristics}

### Data Structure {#sec-dataCharacteristicsStructure}

```{r}
str(nfl_players)
```

### Data Dimensions {#sec-dataCharacteristicsDimensions}

Number of rows and columns:

```{r}
dim(nfl_players)
```

### Number of Elements {#sec-dataCharacteristicsLength}

```{r}
length(nfl_players$display_name)
```

### Number of Missing Elements {#sec-dataCharacteristicsMissingElements}

```{r}
length(nfl_players$college_name[which(is.na(nfl_players$college_name))])
```

### Number of Non-Missing Elements {#sec-dataCharacteristicsNonMissingElements}

```{r}
length(nfl_players$college_name[which(!is.na(nfl_players$college_name))])
length(na.omit(nfl_players$college_name))
```

## Create New Variables {#sec-createNewVars}

To create a new variable, use the following syntax:

```{r}
nfl_players$newVar <- NA
```

Here is an example of creating a new variable:

```{r}
nfl_players$newVar <- 1:nrow(nfl_players)
```

nfl_players$newVar <- NULL
```

## Create a Data Frame {#sec-createDF}

Here is an example of creating a data frame:

```{r}
mydata <- data.frame(
  ID = c(1:5, 1047:1051),
  cat = sample(
    0:1,
    10,
    replace = TRUE)
)

mydata
```

## Recode Variables {#sec-recodeVars}

Here is an example of recoding a variable:

```{r}
mydata$oldVar1 <- NA
mydata$oldVar1[which(mydata$cat == 0)] <- "dog"
mydata$oldVar1[which(mydata$cat == 1)] <- "cat"

mydata$oldVar2 <- NA
mydata$oldVar2[which(mydata$cat == 0)] <- "no"
mydata$oldVar2[which(mydata$cat == 1)] <- "yes"
```

Recode multiple variables:

```{r}
mydata %>%
  mutate(across(c(
    oldVar1:oldVar2),
    ~ case_match(
      .,
      c("dog","no") ~ 0,
      c("cat","yes") ~ 1)))
```

## Rename Variables {#renameVars}

```{r}
mydata <- mydata %>% 
  rename(
    newVar1 = oldVar1,
    newVar2 = oldVar2)
```

Using a vector of variable names:

varNamesFrom <- c("oldVar1","oldVar2")
varNamesTo <- c("newVar1","newVar2")

mydata <- mydata %>% 
  rename_with(~ varNamesTo, all_of(varNamesFrom))
```

## Convert the Types of Variables {#convertVarTypes}

One variable:

```{r}
mydata$factorVar <- factor(mydata$ID)
mydata$numericVar <- as.numeric(mydata$cat)
mydata$integerVar <- as.integer(mydata$cat)
mydata$characterVar <- as.character(mydata$newVar1)
```

Multiple variables:

```{r}
mydata %>%
  mutate(across(c(
    ID,
    cat),
    as.numeric))

mydata %>%
  mutate(across(
    ID:cat,
    as.numeric))

mydata %>%
  mutate(across(where(is.factor), as.character))
```
